Hypertension and microbiome

### 2019-05-06 articletwo.Rmd
Basic PCoA functionality

```{r pca calculate}
if (!file.exists("rds/pca.prcomp.rds")) {
    pca.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
    pca.prcomp <- stats::prcomp(pca.abundances)
    saveRDS(pca.prcomp, file = "rds/pca.prcomp.rds")
} else {
    pca.prcomp <- readRDS("rds/pca.prcomp.rds")
}
```

@@

## Principal component analysis (Figure 2.)

```{r pca plot}
pca.eigs <- pca.prcomp$sdev^2
pca.percentage <- 100 * pca.eigs /sum(pca.eigs)
pca.labels <- sprintf("%s (%.2f %%)", colnames(pca.prcomp$x), pca.percentage)

pcaplot <- ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
    geom_point(aes(color = HYPERTENSION), size = 0.5) +
    scale_color_manual(values=c("blue", "red"),
                       name= "Hypertension",
                       breaks=c("1", "0"),
                       labels=c("Yes", "No")) +
    xlab(pca.labels[1]) +
    ylab(pca.labels[2]) +
    coord_fixed() +
    scale_x_continuous(breaks=seq(-100, 100, 20)) +
    scale_y_continuous(breaks=seq(-30, 30, 10)) +
    theme_classic() +
    theme(text = element_text(size=20),
          legend.justification=c(1,0),
          legend.position=c(1.03, 0.75),
          legend.title=element_text(size = 20),
          legend.direction="vertical",
          legend.text=element_text(size = floor(0.75*20)),
          legend.box.margin = margin(1, 1, 1, 1, "mm"),
          legend.box.background = element_rect(colour = "black"))

ggsave(file = "cache/pca.png", plot = pcaplot, height = 5, width = 9, dpi = 300)
```

>>>>

```{r pcoa calculate}
if (!file.exists("rds/pcoa.ordinate.rds")) {
    pcoa.abundances <- microbiome::transform(pseq.species, 'compositional')
    pcoa.ordinate <- ordinate(pcoa.abundances, method="PCoA", distance="bray")
    saveRDS(pca.prcomp, file = "rds/pcoa.ordinate.rds")
} else {
    pcoa.ordinate <- readRDS("rds/pcoa.ordinate.rds")
}
```

@@

## Principal coordinate analysis (Figure 2.)

```{r pcoa plot}
pcoa.vectors <- pcoa.ordinate$vectors %>% as.data.frame
eig.fracs <- pcoa.ordinate$values$Relative_eig
pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2),
                 meta(pseq.species), by=0, all = TRUE)

colnames(pcoa.df)

pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = HYPERTENSION, shape = ANYDRUG)) +
    geom_point(size=0.4, alpha=0.4) +
    xlab(paste("PCoA axis 1,", round(100*eig.fracs[1],1), "%")) +
    ylab(paste("PCoA axis 2,", round(100*eig.fracs[2],1), "%")) +
    scale_color_manual(values=c("blue", "red"),
                       name= "Hypertension",
                       breaks=c("1", "0"),
                       labels=c("Yes", "No")) +
    scale_shape_manual(values=c(4, 1),
                       name = "BP-medication",
                       breaks = c("1", "0"),
                       labels = c("Yes", "No")) +
    coord_fixed() +
    scale_x_continuous(breaks=seq(-100, 100, 20)) +
    scale_y_continuous(breaks=seq(-30, 30, 10)) +
    theme_classic() +
    guides(colour = guide_legend(override.aes = list(size=6, shape = 15)),
           shape = guide_legend(override.aes = list(size=6))) +
    theme(text = element_text(size=14),
          legend.title = element_text(size = 12),
          legend.text=element_text(size=10),
          legend.justification=c(1,0),
          legend.position=c(1,0),
          legend.margin=margin(t = 0, unit='cm'),
          legend.background = element_rect(fill=alpha('gray', 0.1)))

ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=5, width=7, dpi = 300)

```
